---
title: "Adherence Aim"
format: docx
editor: visual
editor_options: 
  chunk_output_type: console
execute:
  cache: true  
---

## Medication use during a pandemic

Chrystelle Kiang

```{r}
#| include: false
#| label: setup

library(here) # makes file managment easier
library(tidyverse) # data wrangling 
library(AdhereR) # adherence package
library(gtsummary) # creates nice tables
library(patchwork) # plot arranging if we don't do overlaid plots
library(flextable) # to print tables from gtsummary
library(viridis) # for colors
library(RColorBrewer) # more colors

# this dataset has all eligible study population 
# created in data_creation.R 
# this version has multiple tumors per StudyID
pharmstudydata <- readRDS("studypopdata.rds")

# one row per StudyID
elig_pop <- pharmstudydata %>%
  distinct(StudyID, .keep_all = TRUE) 

# subset pts who have rx and those who have rx after diagnosis
rx_before_dx <- pharmstudydata %>%
  filter(dispense_dt <= dt_diagnosis) %>%
    distinct(StudyID, .keep_all = TRUE)

# number with prescriptions outside of study period
rx_before_study <- pharmstudydata %>%
  anti_join(rx_before_dx, by = "StudyID") %>%
  group_by(StudyID) %>%
  arrange(dispense_dt) %>%
   summarise(last_dispense = max(dispense_dt)) %>%
  select(StudyID, last_dispense) %>%
  filter(!is.na(last_dispense),
         last_dispense < as.Date("2017-09-01", format = "%Y-%m-%d")) 

pop_no_rx <- pharmstudydata %>% 
  filter(is.na(dispense_dt)) %>%
      distinct(StudyID, .keep_all = TRUE)

one_rx <- pharmstudydata %>%
  anti_join(pop_no_rx, by = "StudyID") %>% # remove no pharmacy claims
  anti_join(rx_before_study, by = "StudyID") %>%
  anti_join(rx_before_dx, by = "StudyID") %>%
  group_by(StudyID) %>%
  summarise(n_claims = n()) %>%
  ungroup() %>%
  filter(n_claims == 1)

# final study population
study_pop <- pharmstudydata %>% 
  anti_join(pop_no_rx, by = "StudyID") %>% # no pharmacy claims
  anti_join(rx_before_study, by = "StudyID") %>% # claim before study
  anti_join(rx_before_dx, by = "StudyID") %>% # claim before diagnosis
  anti_join(one_rx, by = "StudyID") %>% # not really users
  mutate(daily_dose = 1) 

study_IDs <- study_pop %>%
  distinct(StudyID, .keep_all = TRUE) 
# rx population not included in study
excl_IDs <- elig_pop %>%
  anti_join(pop_no_rx, by = "StudyID") %>% # no pharmacy claims
  anti_join(study_IDs, by = "StudyID")
  
# only keep those with pharm info and after Sept 2017 to allow for fills that occur before 2018-01-01
# this dataset has everyone with and without prescriptions to allow for comparison of who we have claims data on 
pharmstudyIDs <- pharmstudydata %>%
    distinct(StudyID, .keep_all = TRUE) %>%
  mutate(final_study = ifelse(StudyID %in% study_pop$StudyID, 1, 0),
         rx_excluded = ifelse(StudyID %in% excl_IDs$StudyID, 1, 0),
         # 1 for those who were in pharm data
         all_rx = ifelse(StudyID %in% pop_no_rx$StudyID, 0, 1),
          status = case_when(
           final_study == 1 ~ 1,
           rx_excluded == 1 ~ 2,
          final_study == 0 & rx_excluded == 0 ~ 3),
         status = factor(status, 
                         levels = c(1,2,3),
                         labels = c("Study population", "Excluded ineligible claims", "Excluded no claims")))

pop_info <- list(total_elig = n_distinct(elig_pop$StudyID),
                 no_rx = n_distinct(pop_no_rx$StudyID),
                 rx_predx = n_distinct(rx_before_dx$StudyID),
                 rx_prestudy = n_distinct(rx_before_study$StudyID),
                 one_rx = n_distinct(one_rx$StudyID),
                 total_study = n_distinct(study_pop$StudyID),
                 no_rx_percent = n_distinct(pop_no_rx$StudyID)/n_distinct(elig_pop$StudyID))
```

### Study Population

There were `r formatC(pop_info$total_elig, big.mark = ",")` patients diagnosed with ER positive breast cancer at stages I to III between January 1, 2015 to July 31, 2019 and captured in the Georgia Cancer Registry. We did not have pharmacy claims data for `r formatC(pop_info$no_rx, big.mark = ",")` (`r formatC(pop_info$no_rx_percent*100, digits = 4, flag = "#")`%) patients. Among those with pharmacy claims data, `r formatC(pop_info$rx_predx, big.mark = ",")` had endocrine therapy before their date of diagnosis and `r formatC(pop_info$rx_prestudy, big.mark = ",")` only had prescriptions before the study period and were excluded. Our final study population consisted of `r formatC(pop_info$total_study, big.mark = ",")` patients.

```{r}
#| echo: false
#| warning: false
#| label: study-pop-demographic

# subset by demographic variable of interest
demo <- study_pop %>%
  group_by(StudyID) %>%
  slice(n()) %>%
  select(StudyID, CTCDATE_OF_DIAGNOSIS_YYYY, ERstatus, estrogen_receptor, stage_summary, drug_group, sex:reporting_source, urb_rural) %>%
  ungroup()
# list of demographics for inline reporting
demo_list <- list(
  n_asianPI = n_distinct(demo %>% filter(race_ethnicity == "NH API")),
  n_blackAA = n_distinct(demo %>% filter(race_ethnicity == "NH Black")),
  n_hispanic = n_distinct(demo %>% filter(race_ethnicity == "Hispanic")),
  n_AIAN = n_distinct(demo %>% filter(race_ethnicity == "NH AI/AN")),
  n_unknown = n_distinct(demo %>% filter(race_ethnicity == "Unknown by patient")),
  n_white = n_distinct(demo %>% filter(race_ethnicity == "NH White")),
  n_other = n_distinct(demo %>% filter(race_ethnicity %in% c("NH Other"))),
  n_urban = n_distinct(demo %>% filter(urb_rural == "Urban")),
  n_rural = n_distinct(demo %>% filter(urb_rural == "Rural"))
)
```

### Results

Of the `r formatC(pop_info$total_elig, big.mark = ",")` eligible patients diagnosed with ER+ breast cancer between 2015 to 2019, Stage I–III, there were `r formatC(pop_info$total_study, big.mark = ",")` with at least two prescriptions for AI or tamoxifen between January 1, 2018 and December 31, 2021 from CVS or Walgreens. Descriptive statistics of the study population included and those eligible but excluded are presented in Table 2. Overall, there was a decrease in the proportion of patients who were adherent over time from January 1, 2018 to December 31, 2022, ranging between **XXX**. The actual count and proportion of patients who were adherent at each interval and count of those who were not included are in Table 3. 

```{r}
#| echo: false
#| warning: false
#| label: simple-table-1
#| tbl-cap: "Table 1" 

table1 <- study_pop %>% 
  distinct(StudyID, .keep_all = TRUE) %>%
  tbl_summary(include = c(CTCAGE_AT_DIAGNOSIS, stage_summary, estrogen_receptor, laterality, reporting_source, CTCDATE_OF_DIAGNOSIS_YYYY, race_ethnicity, marital_status, urb_rural))
as_flex_table(table1)
```

Also looking at included/excluded patients.
```{r}
#| echo: false
#| warning: false
#| label: table-1s
#| tbl-cap: "Demo tables"

table1_full <- pharmstudyIDs %>% 
  ungroup() %>%
  select(CTCAGE_AT_DIAGNOSIS, stage_summary, estrogen_receptor, laterality, reporting_source, CTCDATE_OF_DIAGNOSIS_YYYY, race_ethnicity, marital_status, urb_rural, status) %>%
  tbl_summary(by = status)

as_flex_table(table1_full)
```

```{r}
#| eval: false
#| echo: false
#| label: run-cma

# this is code that used to get CMA
# first is the function does OW per person then there are a few steps to prep data for function, which is below function 
# function to run CMA, input is dataset and choice of CMA 
run_CMA <- function(data, CMA_choice){
  # creating these so can change inside of function to be able to loop thru
  FU_start <- as.Date("2017-10-09", format = "%Y-%m-%d") 
  FU_end <- as.Date("2022-01-01", format = "%Y-%m-%d")
  FU_duration <- as.integer(difftime(FU_end, FU_start, units = "weeks"))
  Jan1 <- as.Date("2018-01-01", format = "%Y-%m-%d")
  
  OW_duration <- min(208, first(data$OW_duration), na.rm = TRUE)
  
    cma <- CMA_sliding_window(CMA.to.apply = CMA_choice,
                     data = data,
                     ID.colname = "StudyID",
                     event.date.colname = "dispense_dt",
                     event.duration.colname = "days_supply",
                     event.daily.dose.colname ="daily_dose",
                     medication.class.colname = "AETdrug",
                     carry.only.for.same.medication = FALSE,
                     consider.dosage.change = FALSE,
                     followup.window.start = FU_start,
                     followup.window.duration = FU_duration,
                     followup.window.duration.unit = "weeks",
                     observation.window.start = Jan1,
                     observation.window.duration = OW_duration,
                     observation.window.duration.unit = "weeks",
                     sliding.window.start = 0, 
                     sliding.window.start.unit = "weeks",
                     sliding.window.duration = 4,
                     sliding.window.duration.unit = "weeks",
                     # if window and step are same, there is no overlap
                     sliding.window.step.duration = 4,
                     sliding.window.step.unit = "weeks",
                     date.format = "%Y/%m/%d",
                     parallel.backend = "none",
                     parallel.threads = 1) 
  result <- as_tibble(cma$CMA)
  return(result)
}

# Steps I take to make each OW flexible per person
# 1. getting last dispense date for each ID
ID_dispense <- study_pop %>%
  group_by(StudyID) %>%
    arrange(dispense_dt) %>%
     summarise(first_dispense = min(dispense_dt),
          last_dispense = max(dispense_dt),
          days_supply = last(days_supply)) %>%
    mutate(end_supply = last_dispense + days_supply + days(30)) %>%
  select(StudyID, first_dispense, last_dispense, end_supply)

# 2. add columns back
study_run <- study_pop %>%
  inner_join(ID_dispense, by = "StudyID") %>%
  mutate(horm_mo = case_when(is.na(CTCRX_SUMM_DATE_HORMONE_MM) ~ 1,
                              TRUE ~ CTCRX_SUMM_DATE_HORMONE_MM),
         horm_dd = case_when(is.na(CTCRX_SUMM_DATE_HORMONE_DD) ~ 1,
                            TRUE ~ CTCRX_SUMM_DATE_HORMONE_DD),
         horm_dt = case_when(!is.na(CTCRX_SUMM_DATE_HORMONE_YYYY) ~ as.Date(
           paste(CTCRX_SUMM_DATE_HORMONE_YYYY,
                 sprintf("%02d", as.numeric(horm_mo)),
                 sprintf("%02d", as.numeric(horm_dd)),
                 sep = "-"), format = "%Y-%m-%d")),
         horm_end_dt = horm_dt + years(5),
         death_dt = case_when(PATIENTVITAL_STATUS == 0 ~ as.Date(paste(PATIENTDATE_OF_LAST_CONTACT_YYYY, PATIENTDATE_OF_LAST_CONTACT_MM, PATIENTDATE_OF_LAST_CONTACT_DD,
                 sep = "-"), format = "%Y-%m-%d")),
         OW_end = case_when(PATIENTVITAL_STATUS == 0 ~ max(as.Date("2018-01-01",format = "%Y-%m-%d"), death_dt),
                            PATIENTVITAL_STATUS == 1 ~ max(end_supply, horm_end_dt, na.rm = TRUE)),
         OW_duration = as.integer(ceiling((difftime(OW_end, "2018-01-01", units = "weeks"))/4) * 4),
         OW_duration = ifelse(OW_duration < 0, 0, OW_duration))

# 3. loop through
result_list <- c()
for (i in unique(study_run$StudyID)){
  study_data <- study_run[study_run$StudyID == i, ]
  result <- run_CMA(study_data, "CMA7")
  result_list[[i]] <- result
}

cma_results <- bind_rows(result_list)

# then can use get_cma_results function to get the proportion adherent and etc
# or just code to summarize 
```

```{r}
#| include: false
#| label: get-results-function

# function that outputs regression table and basic ggplot
# input dataset is the output of CMA results saved as tibble
get_cma_results <- function(dataset, COVID_start){
  reg_output <- c()
  reg_summary <- c()
  reg_plot <- c()
  results_list <- c()
  
  dataset <- dataset %>%
      mutate(adherent = case_when(
      CMA >= 0.80 ~ 1,
      CMA < 0.80 ~ 0)) %>%
    group_by(window.start) %>%
    reframe(mean_CMA = mean(CMA, na.rm = TRUE),
            adherent_count = sum(adherent == 1, na.rm = TRUE), 
            non_adherent_count = sum(adherent == 0, na.rm = TRUE),
            adherent_NA_count = sum(is.na(adherent)),
            prop_adherent = mean(adherent, na.rm = TRUE),
            prop_NA = adherent_NA_count/(adherent_NA_count + adherent_count + adherent_NA_count)) %>%
    arrange(window.start) %>%
    mutate(post_COVID = if_else(window.start >= COVID_start, 1, 0),
           window.ID = row_number()) %>%
    ungroup()
  
  its <- lm(prop_adherent ~ window.ID + post_COVID + window.ID*post_COVID, data = dataset)
  reg_summary <- summary(its)
  
  dataset2 <- dataset %>% na.omit(prop_adherent) %>%
    mutate(fit = predict(its))

  reg_plot <- ggplot(dataset2, aes(x = window.start, y = prop_adherent)) +
    geom_point() +
    geom_line(aes(y = fit), colour = "#377EB8", linewidth = 1) +
    geom_vline(xintercept = COVID_start, colour = "#E41A1C") 
  
  results_list <- list(data_summary = dataset,
                       plot_data = dataset2,
                       reg_output = its,
                   reg_summary = reg_summary,
                   reg_plot = reg_plot)
  return(results_list)
}
```

Main Analyses

```{r}
#| echo: false
#| warning: false
#| label: main-its
#| fig-cap: "Overall ITS" 

# for the sake of time, loading results from before
# but were done using function above
# run March 26 in notebook file and filtered for IDs with 2+ claims
cma7 <- readRDS("./data/cma7.rds")

march14 <- as.Date("2020-03-14", format = "%Y-%m-%d")
cma7_results <- get_cma_results(cma7, march14)
# function output is dataset with summaries for each time interval, regression model, regression summary, and plot for times 
cma7_data <- cma7_results$data_summary
cma7_plot_data <- cma7_results$plot_data

# write_csv(cma7_data, "./data/cma7-table.csv")
main_plot <- cma7_results$reg_plot + 
labs(x = "Year", y = "Proportion Adherent (80%)") +
  scale_y_continuous(limits = c(0.1,1)) + 
  theme_bw()
# main_plot

ggplot(cma7_plot_data, aes(x = window.start, y = prop_adherent)) +
    geom_point(size = 2) + # geom_line(aes(y = prop_adherent), linetype = "dotted", colour = "black", alpha = 0.7, linewidth = 1) +
    geom_line(aes(y = fit), colour = "#377EB8", linewidth = 1) +
    geom_vline(xintercept = march14, colour = "#E41A1C") +  
  scale_y_continuous(limits = c(0.1,1)) + 
  theme_bw() + 
  labs(x = "Year", y = "Proportion Adherent (80%)") 

```

In our ITS analyses, the proportion of adherent patients decreased by 0.51% (95% CI -0.56, -0.46) every 4 weeks from January 1, 2018 to the onset of the pandemic in March 14, 2020. At pandemic onset, the linear proportion adherent increased by 4.3% (95% CI 1.25%, 7.38%). During the pandemic, the proportion adherence remained mostly unchanged, continuing to decrease over time but with a slight increase in slope of 0.04% (95% CI -0.04%, 0.05%) in the pandemic years relative to pre-pandemic years. Regression estimates are in Table 4 and Figure 4 is a plot of the proportion of adherence over time and the fitted linear ITS regression line. Note that the change in actual proportion in the interval right at pandemic onset (April 23, 2020 to March 24, 2020) differs from the ITS estimates, which are based on the difference in the regression lines of the whole pre- and post-pandemic study period.

```{r}
#| echo: false
#| warning: false
#| label: main-reg-table
#| tbl-cap: "Regression table" 


as_flex_table(cma7_results$reg_summary %>%
  tbl_regression(estimate_fun = ~ style_number(.x, digits = 4)))
```

#### Subgroup analyses

We repeated this among demographic subgroups of interest, by race/ethnicity, county-level urbanicity (at time of diagnosis), marital status (at time of diagnosis), and year of diagnosis.

##### Race/ethnicity

There are `r formatC(demo_list$n_white, big.mark = ",")` patients of White race, `r formatC(demo_list$n_blackAA, big.mark = ",")` Black, `r formatC(demo_list$n_asianPI, big.mark = ",")` Asian or Pacific Islander, `r formatC(demo_list$n_AIAN, big.mark = ",")` Alaskan Native, American Indian, `r demo_list$n_other` other, and `r demo_list$n_unknown` unknown (to patinet) race, and `r formatC(demo_list$n_hispanic, big.mark = ",")` Hispanic ethnicity.

```{r}
#| echo: false
#| warning: false
#| label: by-race
#| fig-cap: "By race" 

# NH White 
# 1. subset CMAs to that race/ethnicity group
pdc_white <- demo %>%
  filter(race_ethnicity == "NH White") %>%
  left_join(cma7, by = "StudyID") %>%
  get_cma_results(march14)
# 2. get_cma_results function outputs a list, plot_data has data + fit
pdc_white_data <- pdc_white$plot_data %>% 
# 3. creating label variable for later overlaying 
  mutate(race_group = "NH White")

# NH Black or African American
pdc_blackAA <- demo %>%
  filter(race_ethnicity == "NH Black") %>%
  left_join(cma7, by = "StudyID") %>%
  get_cma_results(march14)
pdc_black_data <- pdc_blackAA$plot_data %>% 
  mutate(race_group = "NH Black")

# Asian/PI
pdc_API <- demo %>%
  filter(race_ethnicity == "NH API") %>%
  left_join(cma7, by = "StudyID") %>%
  get_cma_results(march14)
pdc_API_data <- pdc_API$plot_data %>%
  mutate(race_group = "NH API")

# Hispanic
pdc_hisp <- demo %>%
  filter(race_ethnicity == "Hispanic") %>%
  left_join(cma7, by = "StudyID") %>%
  get_cma_results(march14)
pdc_hisp_data <- pdc_hisp$plot_data %>%
  mutate(race_group = "Hispanic")

pdc_merged_race <- rbind(pdc_white_data, pdc_black_data, pdc_API_data, pdc_hisp_data)

# colors from Viridis Color Palatte Generator
plasma5 <- c("#f89540", "#cc4778", "#7e03a8", "#0d0887")

plot_by_race <- ggplot(pdc_merged_race, aes(x = window.start, y = prop_adherent, color = race_group)) +
    geom_point() + 
    geom_line(aes(y = fit, group = race_group)) +
    geom_vline(xintercept = march14, colour = "red") +  
  scale_y_continuous(limits = c(0.1,1)) + 
  theme_bw() + scale_color_manual(values = plasma5) +
  labs(x = "Year", y = "Proportion Adherent (80%)") 
  #scale_color_viridis(discrete=TRUE, option="magma") 
# plot_by_race 

plot_by_race_scatter <- ggplot(pdc_merged_race, aes(x = window.start, y = prop_adherent, color = race_group)) +
    geom_point() + geom_line(linetype = "dotted") +
    geom_vline(xintercept = march14, colour = "red") +  
  scale_y_continuous(limits = c(0.1,1)) + 
  theme_bw() + scale_color_manual(values = plasma5) + 
  theme(legend.title=element_blank()) +
  labs(x = "Year", y = "Proportion Adherent (80%)") 
  #scale_color_viridis(discrete=TRUE, option="magma") 
plot_by_race_its <- 
  ggplot(pdc_merged_race, aes(x = window.start, y = fit, color = race_group)) +
    geom_line(linewidth = 1) +
    geom_vline(xintercept = march14, colour = "red") +  
  scale_y_continuous(limits = c(0.1,1)) + 
  theme_bw() + scale_color_manual(values = plasma5) +
  theme(legend.title=element_blank()) +
  labs(x = "Year", y = "Proportion Adherent (80%)") 

race_plots <- (plot_by_race_scatter/plot_by_race_its) + plot_annotation(tag_levels = 'A')
race_plots
```

By race/ethnicity, adherence appears to have greater range among NH Black or AA, NH Asian, and Hispanic patients compared to NH White patients.

```{r}
#| echo: false
#| warning: false
#| label: subgroup-tables
#| tbl-cap: "Regressions by race"

as_flex_table(pdc_white$reg_summary %>%
  tbl_regression(estimate_fun = ~ style_number(.x, digits = 4)) %>%
  modify_spanning_header(c(estimate, ci, p.value) ~ "**NH White**"))

as_flex_table(pdc_blackAA$reg_summary %>%
  tbl_regression(estimate_fun = ~ style_number(.x, digits = 4)) %>%
  modify_spanning_header(c(estimate, ci, p.value) ~ "**NH Black**"))
  
as_flex_table(pdc_API$reg_summary %>%
  tbl_regression(estimate_fun = ~ style_number(.x, digits = 4)) %>%
  modify_spanning_header(c(estimate, ci, p.value) ~ "**NH API**"))

as_flex_table(pdc_hisp$reg_summary %>%
  tbl_regression(estimate_fun = ~ style_number(.x, digits = 4)) %>%
  modify_spanning_header(c(estimate, ci, p.value) ~ "**Hispanic**")) 
```

##### Rural/urban

There are `r formatC(demo_list$n_urban, big.mark = ",")` patients who resided in urban or suburban counties, defined as XXX,XXX people, and `r formatC(demo_list$n_rural, big.mark = ",")` patients in rural counties.

```{r}
#| echo: false
#| warning: false
#| label: urban-rural
#| fig-cap: "Urban and Rural" 
#| tbl-cap: "Urban and Rural" 

# Urban 
pdc_urban <- demo %>%
  filter(urb_rural == "Urban") %>%
  left_join(cma7, by = "StudyID") %>%
  get_cma_results(march14)
pdc_urban_data <- pdc_urban$plot_data %>%
  mutate(county = "Urban")
# Rural
pdc_rural <- demo %>%
  filter(urb_rural == "Rural") %>%
  left_join(cma7, by = "StudyID") %>%
  get_cma_results(march14)
pdc_rural_data <- pdc_rural$plot_data %>%
  mutate(county = "Rural")

pdc_counties <- rbind(pdc_urban_data, pdc_rural_data)
county_colors <- c("#bae4bc","#0868ac")
plot_county_scatter <- 
  ggplot(pdc_counties, aes(x = window.start, y = prop_adherent, color = county)) +
    geom_point() + geom_line(linetype = "dotted") +
    geom_vline(xintercept = march14, colour = "red") +  
  scale_y_continuous(limits = c(0.1,1)) + 
  theme_bw() + scale_color_manual(values = county_colors) +
  labs(x = "Year", y = "Proportion Adherent (80%)") +
  theme(legend.title=element_blank()) 

plot_county_its <- 
  ggplot(pdc_counties, aes(x = window.start, y = fit, color = county)) +
    geom_line(linewidth = 1) +
    geom_vline(xintercept = march14, colour = "red") +  
  scale_y_continuous(limits = c(0.1,1)) + 
  theme_bw() + scale_color_manual(values = county_colors) +
  labs(x = "Year", y = "Proportion Adherent (80%)") +
  theme(legend.title=element_blank())

county_plots <- (plot_county_scatter/plot_county_its) + plot_annotation(tag_levels = 'A')
county_plots

as_flex_table(pdc_urban$reg_summary %>%
 tbl_regression(estimate_fun = ~ style_number(.x, digits = 4))%>%
  modify_spanning_header(c(estimate, ci, p.value) ~ "**Urban**"))

as_flex_table(pdc_rural$reg_summary %>%
  tbl_regression(estimate_fun = ~ style_number(.x, digits = 4)) %>%
  modify_spanning_header(c(estimate, ci, p.value) ~ "**Rural**"))
```

##### Marital status

```{r}
#| echo: false
#| warning: false
#| label: marital-stat
#| fig-cap: "By Marital Status"
#| tbl-cap: "Marital Status"

# Single (never married) 
pdc_single <- demo %>%
  filter(marital_status == "Single (never married)") %>%
  left_join(cma7, by = "StudyID") %>%
  get_cma_results(march14)
pdc_single_data <- pdc_single$plot_data %>%
  mutate(status = "Single")
# Married/partnered
pdc_married <- demo %>%
  filter(marital_status %in% c("Married (including common law)", "Unmarried or Domestic Partner")) %>%
  left_join(cma7, by = "StudyID") %>%
  get_cma_results(march14)
pdc_married_data <- pdc_married$plot_data %>%
  mutate(status = "Married")
# No longer married
pdc_unmarried <- demo %>%
  filter(marital_status %in% c("Separated", "Divorced", "Widowed")) %>%
  left_join(cma7, by = "StudyID") %>%
  get_cma_results(march14)
pdc_unmarried_data <- pdc_unmarried$plot_data %>%
  mutate(status = "No longer married")

pdc_marital <- rbind(pdc_single_data, pdc_married_data, pdc_unmarried_data)

plot_marital_scatter <- ggplot(pdc_marital, aes(x = window.start, y = prop_adherent, color = status)) +
    geom_point() + geom_line(linetype = "dotted") +
    geom_vline(xintercept = march14, colour = "red") +  
  scale_y_continuous(limits = c(0.1,1)) + 
  theme_bw() + scale_color_brewer(palette = "Set1") +
  theme(legend.title=element_blank()) +
  labs(x = "Year", y = "Proportion Adherent (80%)") 

plot_marital_its <- ggplot(pdc_marital, aes(x = window.start, y = fit, color = status)) +
    geom_line(linewidth = 1) +
    geom_vline(xintercept = march14, colour = "red") +  
  scale_y_continuous(limits = c(0.1,1)) + 
  theme_bw() + scale_color_brewer(palette = "Set1") +
  theme(legend.title=element_blank()) +
  labs(x = "Year", y = "Proportion Adherent (80%)") 

marital_plots <- (plot_marital_scatter/plot_marital_its) + plot_annotation(tag_levels = 'A')
marital_plots

as_flex_table(pdc_single$reg_summary %>%
 tbl_regression(estimate_fun = ~ style_number(.x, digits = 4))%>%
  modify_spanning_header(c(estimate, ci, p.value) ~ "**Single**"))

as_flex_table(pdc_married$reg_summary %>%
  tbl_regression(estimate_fun = ~ style_number(.x, digits = 4)) %>%
  modify_spanning_header(c(estimate, ci, p.value) ~ "**Married**"))

as_flex_table(pdc_unmarried$reg_summary %>%
  tbl_regression(estimate_fun = ~ style_number(.x, digits = 4)) %>%
  modify_spanning_header(c(estimate, ci, p.value) ~ "**Unmarried**"))
```

##### Diagnosis year

```{r}
#| echo: false
#| warning: false
#| label: dx-year
#| fig-cap: "Diagnosis year" 
#| tbl-cap: "Dx year"

study_start <- as.Date("2018-01-01", format = "%Y-%m-%d")
study_end <- as.Date("2022-01-01", format = "%Y-%m-%d")

# 2015
pdc_2015 <- demo %>%
  filter(CTCDATE_OF_DIAGNOSIS_YYYY == 2015) %>%
  left_join(cma7, by = "StudyID") %>%
  get_cma_results(march14)
pdc_dx2015 <- pdc_2015$plot_data %>%
  mutate(yeardx = "2015")
 
pdc_2016 <- demo %>%
  filter(CTCDATE_OF_DIAGNOSIS_YYYY == 2016) %>%
  left_join(cma7, by = "StudyID") %>%
  get_cma_results(march14)
pdc_dx2016 <- pdc_2016$plot_data %>%
  mutate(yeardx = "2016")

pdc_2017 <- demo %>%
  filter(CTCDATE_OF_DIAGNOSIS_YYYY == 2017) %>%
  left_join(cma7, by = "StudyID") %>%
  get_cma_results(march14)
pdc_dx2017 <- pdc_2017$plot_data %>%
  mutate(yeardx = "2017")
 
pdc_2018 <- demo %>%
  filter(CTCDATE_OF_DIAGNOSIS_YYYY == 2018) %>%
  left_join(cma7, by = "StudyID") %>%
  get_cma_results(march14)
pdc_dx2018 <- pdc_2018$plot_data %>%
  mutate(yeardx = "2018")

# for dx year 2019, have no estimates for 2018 and that was making the regression wonky so only doing times after 2019
jan2019 <- as.Date("2019-01-01", format = "%Y-%m-%d")
pdc_2019 <- demo %>%
  filter(CTCDATE_OF_DIAGNOSIS_YYYY == 2019) %>%
  left_join(cma7, by = "StudyID") %>%
  filter(window.start >= jan2019) %>%
      mutate(adherent = case_when(
      CMA >= 0.80 ~ 1,
      CMA < 0.80 ~ 0)) %>%
    group_by(window.start) %>%
    reframe(mean_CMA = mean(CMA, na.rm = TRUE),
            adherent_count = sum(adherent == 1, na.rm = TRUE), 
            non_adherent_count = sum(adherent == 0, na.rm = TRUE),
            adherent_NA_count = sum(is.na(adherent)),
            prop_adherent = mean(adherent, na.rm = TRUE),
            prop_NA = adherent_NA_count/(adherent_NA_count + adherent_count + adherent_NA_count)) %>%
    arrange(window.start) %>%
    mutate(post_COVID = if_else(window.start >= march14, 1, 0),
           window.ID = row_number()) %>%
    ungroup()

dx2019_its <- lm(prop_adherent ~ window.ID + post_COVID + window.ID*post_COVID, data = pdc_2019)
dx2019_reg_summary <- summary(dx2019_its)
# now making it match others  
pdc_dx2019 <- pdc_2019 %>% 
    mutate(fit = predict(dx2019_its),
           yeardx = "2019")

pdc_by_year <- rbind(pdc_dx2015, pdc_dx2016, pdc_dx2017, pdc_dx2018, pdc_dx2019)

plot_year_scatter <- ggplot(pdc_by_year, aes(x = window.start, y = prop_adherent, color = yeardx)) +
    geom_point() + geom_line(linetype = "dotted") +
    geom_vline(xintercept = march14, colour = "red") +  
  scale_y_continuous(limits = c(0.1,1)) + 
  theme_bw() +  
  labs(x = "Year", y = "Proportion Adherent (80%)") +
  theme(legend.title=element_blank()) +
  scale_color_viridis(discrete=TRUE, option="viridis") 
plot_year_its <- 
  ggplot(pdc_by_year, aes(x = window.start, y = fit, color = yeardx)) +
    geom_line(linewidth = 1) +
    geom_vline(xintercept = march14, colour = "red") +  
  scale_y_continuous(limits = c(0.1,1)) + 
  theme_bw() + 
  labs(x = "Year", y = "Proportion Adherent (80%)") + 
  theme(legend.title=element_blank()) +
  scale_color_viridis(discrete=TRUE, option="viridis") 

dx_year_plots <- (plot_year_scatter / plot_year_its) + plot_annotation(tag_levels = 'A')
dx_year_plots

as_flex_table(pdc_2015$reg_summary %>%
  tbl_regression(estimate_fun = ~ style_number(.x, digits = 4)) %>%
  modify_spanning_header(c(estimate, ci, p.value) ~ "**Dx 2015 (Years 4 to 7)**"))
as_flex_table(pdc_2016$reg_summary %>%
  tbl_regression(estimate_fun = ~ style_number(.x, digits = 4)) %>%
  modify_spanning_header(c(estimate, ci, p.value) ~ "**Dx 2016 (Years 3 to 6)**"))
as_flex_table(pdc_2017$reg_summary %>%
  tbl_regression(estimate_fun = ~ style_number(.x, digits = 4)) %>%
  modify_spanning_header(c(estimate, ci, p.value) ~ "**Dx 2017 (Years 2 to 5)**"))
as_flex_table(pdc_2018$reg_summary %>%
  tbl_regression(estimate_fun = ~ style_number(.x, digits = 4)) %>%
  modify_spanning_header(c(estimate, ci, p.value) ~ "**Dx 2018 (Years 1 to 4)**"))
as_flex_table(dx2019_reg_summary %>%
  tbl_regression(estimate_fun = ~ style_number(.x, digits = 4)) %>%
  modify_spanning_header(c(estimate, ci, p.value) ~ "**Dx 2019 (Years 1 to 3)**"))
```

### Additional material

#### Prescriptions & hormone therapy

Recorded therapy dates, number of claims. Investigating this beceuse treatment window (OW) is based on this date.

```{r}
#| echo: false
#| eval: false
#| label: hormone-tx-start

# keeping these variables as done before
disp_data <- study_pop %>%
  group_by(StudyID) %>%
    arrange(dispense_dt) %>%
     summarise(first_dispense = min(dispense_dt),
          last_dispense = max(dispense_dt),
          days_supply = last(days_supply),
          n_claims = n()) %>%
    mutate(end_supply = last_dispense + days_supply + days(30)) %>%
  ungroup()

# merging back 
each_person <- study_pop %>%
  group_by(StudyID) %>%
  slice(n()) %>%
  inner_join(disp_data, by = "StudyID")

each_person <- each_person %>%
  mutate(summ_hormone = factor(CTCRX_SUMM_HORMONE,
                               levels = c(0,1,82,85,86,87,88,99),
                               labels = c("None, not planned first course", 
                                          "Hormone admin as first course",
                                          "Hormone tx not admin bc patient risk factors",
                                          "Hormone tx not admin bc died",
                                          "Hormone tx recommended but not admin",
                                          "Hormone tx recommended but refused",
                                          "Hormone tx was recommended but unk if admin",
                                          "Not in patient record")),
         horm_mo = case_when(is.na(CTCRX_SUMM_DATE_HORMONE_MM) ~ 1,
                              TRUE ~ CTCRX_SUMM_DATE_HORMONE_MM),
         horm_dd = case_when(is.na(CTCRX_SUMM_DATE_HORMONE_DD) ~ 1,
                            TRUE ~ CTCRX_SUMM_DATE_HORMONE_DD),
         horm_dt = case_when(!is.na(CTCRX_SUMM_DATE_HORMONE_YYYY) ~ as.Date(
           paste(CTCRX_SUMM_DATE_HORMONE_YYYY,
                 sprintf("%02d", as.numeric(horm_mo)),
                 sprintf("%02d", as.numeric(horm_dd)),
                 sep = "-"), format = "%Y-%m-%d")),
         horm_end_dt = horm_dt + years(5),
         death_dt = case_when(PATIENTVITAL_STATUS == 0 ~ as.Date(paste(PATIENTDATE_OF_LAST_CONTACT_YYYY, PATIENTDATE_OF_LAST_CONTACT_MM, PATIENTDATE_OF_LAST_CONTACT_DD,
                 sep = "-"), format = "%Y-%m-%d")),
         OW_end_horm = case_when(PATIENTVITAL_STATUS == 0 ~ max(as.Date("2018-01-01",format = "%Y-%m-%d"), death_dt),
                            PATIENTVITAL_STATUS == 1 ~ max(end_supply, horm_end_dt, na.rm = TRUE)),
         # this was set to 2018-01-01 because that's our study period 
         # TODO create another if, where this is case for dx before 2018, else it's date of diagnosis
         OW_duration_horm = as.integer(ceiling((difftime(OW_end_horm, "2018-01-01", units = "weeks"))/4) * 4),
         OW_duration_horm = ifelse(OW_duration_horm < 0, 0, OW_duration_horm)
  )

horm_info <- each_person %>%
  group_by(StudyID) %>%
  summarise(time_to_horm = as.numeric(difftime(horm_dt, dt_diagnosis, units = "weeks")),
            time_to_OWend = as.numeric(difftime(OW_end_horm, dt_diagnosis, units = "weeks")),
            summ_hormone = summ_hormone,
            n_claims = n_claims)

horm_info <- horm_info %>%
  mutate(five_claims = if_else(n_claims > 5, "more than 5", "1 to 5"))

table(horm_info$summ_hormone, horm_info$five_claims)
summary(horm_info$n_claims)
horm_info %>% group_by(summ_hormone) %>%
  summarise(mean_claims = mean(n_claims))

# time between diagnosis and date of hormone therapy start
ggplot(horm_info, aes(time_to_horm)) + 
  geom_histogram()

summary(each_person$horm_dt) 
summary(each_person$CTCRX_SUMM_DT_SURGERY_YYYY)
table(each_person$CTCRX_SUMM_DATE_HORMONE_YYYY, useNA = "always")
table(each_person$CTCRX_SUMM_DATE_HORMONE_YYYY, each_person$CTCRX_SUMM_DT_SURGERY_YYYY, useNA = "always")
table(each_person$CTCRX_SUMM_DATE_HORMONE_YYYY, each_person$CTCDATE_OF_DIAGNOSIS_YYYY, useNA = "always")
```

Have 925 patients missing date of hormone therapy; they get censored at end of supply. It does look like some people are recorded as not initiating hormone therapy or refusing *as first course*, but indeed have claims.

**Duration:** The duration of prescription ("days supply") ranged from 1 to 180 days, with most being either 30 or 90 days.

```{r}
#| echo: false
#| warning: false
#| label: rx-duration
#| tbl-cap: "prescription-duration"
rx_days_irreg_groups <- study_pop %>%
  ungroup() %>%
  mutate(days_supply_group = factor(case_when(
    days_supply %in% 1:29 ~ "1 to 29",
    days_supply == 30 ~ "30",
    days_supply %in% 31:89 ~ "31 to 89",
    days_supply == 90 ~ "90",
    days_supply %in% 91:179 ~ "91 to 179",
    days_supply == 180 ~ "180"), 
    levels = c("1 to 29", "30", "31 to 89", "90", "91 to 179", "180"),
    labels = c("1 to 29", "30", "31 to 89", "90", "91 to 179", "180")
    )) %>%
  select(days_supply_group)

as_flex_table(rx_days_irreg_groups %>% tbl_summary())

rx_days <- rx_days_irreg_groups %>%
  group_by(days_supply_group) %>%
  reframe(days_count = n()) %>%
  mutate(days_percent = days_count/sum(days_count)*100,
         days_p_label = case_when(
           days_percent >= 0.1 ~ paste(round(days_percent,1),"%", sep = ""),
           days_percent < 0.1 ~ "<0.1%"),
         days_label = paste(days_count," (",round(days_percent,2),"%)", sep = "")) 

ggplot(rx_days, aes(x = days_supply_group, y = days_percent)) +
  geom_col() + 
  geom_text(aes(label = days_p_label), stat = 'identity', vjust = -0.5) +
  scale_x_discrete(expand = c(0,0)) + 
  scale_y_continuous(expand = c(0,0), limits = c(0,70), breaks = c(0,20,40,60)) + 
  labs(x = "Days supply", y = "Percent of prescriptions") +
  theme(text = element_text(size = 20),
        axis.text.x = element_text(color = "black", size = 20),
        axis.text.y = element_text(color = "black", size = 20)) +
  theme_classic() 

study_pop %>% group_by(StudyID) %>% summarise(mean(days_supply))
```

**Type of prescription: AI vs tamoxifen vs both**

```{r}
#| echo: false
#| warning: false
#| label: et-type-tables
#| fig-cap: "by ET type"
drug_indicate <- study_pop %>%
  group_by(StudyID) %>%
  reframe(n_drugs = n_distinct(drug_group),
          which_drug = unique(drug_group)) %>%
  mutate(drug_use = case_when(
    n_drugs == 2 ~ "Both",
    n_drugs == 1 & which_drug == "AI" ~ "AI",
    n_drugs == 1 & which_drug == "TAMOXIFEN" ~ "Tamoxifen")
  ) %>%
  select(StudyID, drug_use) %>%
  distinct()
# Both AI and TAM 
bothAItam <- drug_indicate %>%
  filter(drug_use == "Both") %>%
  left_join(cma7, by = "StudyID") %>%
  get_cma_results(march14)
pdc_both <- bothAItam$plot_data %>%
  mutate(user = "Both")
# TAM only
onlyTAM <- drug_indicate %>%
  filter(drug_use == "Tamoxifen") %>%
  left_join(cma7, by = "StudyID") %>%
  get_cma_results(march14)
pdc_tam <- onlyTAM$plot_data %>%
  mutate(user = "Tamoxifen")
# AI only
onlyAI <- drug_indicate %>%
  filter(drug_use == "AI") %>%
  left_join(cma7, by = "StudyID") %>%
  get_cma_results(march14)
pdc_ai <- onlyAI$plot_data %>%
  mutate(user = "AI")

pdc_drug <- rbind(pdc_both, pdc_tam, pdc_ai)

plot_drug_scatter <- 
  ggplot(pdc_drug, aes(x = window.start, y = prop_adherent, color = user)) +
    geom_point() + geom_line(linetype = "dotted") +
    geom_vline(xintercept = march14, colour = "red") +  
  scale_y_continuous(limits = c(0.1,1)) + 
  theme_bw() + scale_color_brewer(palette = "PuRd") +
  labs(x = "Year", y = "Proportion Adherent (80%)") +
  theme(legend.title=element_blank()) 

plot_drug_its <- 
  ggplot(pdc_drug, aes(x = window.start, y = fit, color = user)) +
    geom_line(linewidth = 1) +
    geom_vline(xintercept = march14, colour = "red") +  
  scale_y_continuous(limits = c(0.1,1)) + 
  theme_bw() + scale_color_brewer(palette = "PuRd") +
  labs(x = "Year", y = "Proportion Adherent (80%)") +
  theme(legend.title=element_blank())

drug_plots <- (plot_drug_scatter/plot_drug_its) + plot_annotation(tag_levels = 'A')
drug_plots
```

```{r}
#| echo: false
#| warning: false
#| label: et-type
#| tbl-cap: "by ET type"
as_flex_table(bothAItam$reg_summary %>%
 tbl_regression(estimate_fun = ~ style_number(.x, digits = 4))%>%
  modify_spanning_header(c(estimate, ci, p.value) ~ "**Both**"))

as_flex_table(onlyTAM$reg_summary %>%
  tbl_regression(estimate_fun = ~ style_number(.x, digits = 4)) %>%
  modify_spanning_header(c(estimate, ci, p.value) ~ "**Tamoxifen only**"))

as_flex_table(onlyAI$reg_summary %>%
  tbl_regression(estimate_fun = ~ style_number(.x, digits = 4)) %>%
  modify_spanning_header(c(estimate, ci, p.value) ~ "**AI only**"))
```

#### Start of COVID

```{r}
#| echo: false
#| warning: false
#| label: change-in-COVID
#| fig-cap: "new start"
#| tbl-cap: "new start"

April19 <- as.Date("2020-04-19", format = "%Y-%m-%d")
april_start <- get_cma_results(cma7, April19)
april_start_data <- april_start$plot_data %>%
  mutate(start = "April")

march_start <- cma7_plot_data %>%
  mutate(start = "March")

may20 <- as.Date("2020-05-20", format = "%Y-%m-%d")
may_start <- get_cma_results(cma7, may1)
may_start_data <- may_start$plot_data %>%
  mutate(start = "May")
diff_starts <- rbind(april_start_data, march_start, may_start_data)

april_plot <- ggplot(april_start_data, aes(x = window.start, y = prop_adherent)) +
    geom_point(size = 2) + geom_line(aes(y = prop_adherent), linetype = "dotted", colour = "black", alpha = 0.7, linewidth = 1) +
    geom_line(aes(y = fit), colour = "#22a884", linewidth = 1) +
    geom_vline(xintercept = April19, colour = "#E41A1C") +  
  scale_y_continuous(limits = c(0.1,1)) + 
  theme_bw() + 
  labs(x = "Year", y = "Proportion Adherent (80%)") 

plot_colors <- c("#A0DA39","#377EB8","#46327E")
# plot_lines <- c("solid", "dashed","solid")

ggplot(diff_starts, aes(x = window.start, y = fit, color = start)) +
  geom_point(aes(y = prop_adherent), size = 1, colour = "black") + 
  geom_line(linewidth = 1.5, alpha = 0.8) + #, aes(linetype = start)) +
  scale_color_manual(values = plot_colors) +
  #scale_linetype_manual(values = plot_lines) +
  scale_y_continuous(limits = c(0.1,1)) + 
  theme_bw() + 
  labs(x = "Year", y = "Proportion Adherent (80%)") +
  theme(legend.title=element_blank()) 
```

```{r}
#| echo: false
#| warning: false
#| label: change-in-COVID-indiv
#| fig-cap: "Individual level time zero"

Jan1 <- as.Date("2018-01-01", format = "%Y-%m-%d") 
COVID_start <- as.Date("2020-03-14", format = "%Y-%m-%d")
COVID_week <- as.integer((difftime(COVID_start, Jan1, units ="weeks"))/4)

covid_IDs <- study_pop %>%
  group_by(StudyID) %>%
  filter(dispense_dt >= COVID_start) %>%
  arrange(dispense_dt) %>%
  slice_head(n = 1) %>%
  mutate(post_dispense = dispense_dt,
         post_dispense_end = post_dispense + days_supply,
         post_weekID = as.integer(floor(difftime(post_dispense_end, Jan1, units ="weeks"))/4),
         last_supply = days_supply) %>%
  select(StudyID, post_dispense, post_dispense_end, post_weekID, last_supply)

indiv_start <- cma7 %>%
  left_join(covid_IDs, by = "StudyID") %>%
  mutate(post_weekID = if_else(is.na(post_weekID), COVID_week, post_weekID))

indiv_level <- indiv_start %>%
    mutate(adherent = case_when(
    CMA >= 0.80 ~ 1,
    CMA < 0.80 ~ 0),
    new_window_ID = window.ID - post_weekID) %>% 
    # grouping by new window ID to align everyone's time 0 
  group_by(new_window_ID) %>%
  reframe(adherent_count = sum(adherent == 1, na.rm = TRUE), 
            non_adherent_count = sum(adherent == 0, na.rm = TRUE),
            adherent_NA_count = sum(is.na(adherent)),
            prop_adherent = mean(adherent, na.rm = TRUE),
            prop_NA = adherent_NA_count/(adherent_NA_count + adherent_count + adherent_NA_count),
          inverse_prop_NA = 1 - prop_NA,
          post_COVID = if_else(new_window_ID >= 0, 1, 0)) %>%
  distinct()

its_indiv_level <- lm(prop_adherent ~ new_window_ID + post_COVID + new_window_ID*post_COVID, data = indiv_level)
summary(its_indiv_level)
# have 1 missing

# omitting the time windows that are missing prop_adherent
indiv_level <- indiv_level %>% na.omit(prop_adherent) %>%
  mutate(total_count = adherent_count + non_adherent_count + adherent_NA_count)
  
indiv_level$fit <- predict(its_indiv_level)

plot_COVID_centered <- 
  ggplot(indiv_level, aes(x = new_window_ID, y = prop_adherent)) +
  geom_point() +
  # geom_line(aes(y = fit), colour = "blue", linewidth = 1) +
  geom_vline(xintercept = 0, colour = "#E41A1C") +
  labs(x = "Weeks relative to first post-pandemic prescription end", y = "Proportion Adherent (80%)") +
  scale_y_continuous(limits = c(0.1,1)) + theme_bw()

plot_COVID_centered
```

This is a bit hard to interpret, but overall do see the same trend of an uptick at pandemic and later decline.

#### Other CMAs

Compared different CMAs. CMA7 is based on observation window, which varies between patients, while CMA4, 5, and 6 are based on dispense dates.

```{r}
#| echo: false
#| warning: false
#| label: other-CMAs
#| fig-cap: "CMAs"
#| tbl-cap: "CMAs"

cma6 <- readRDS("./data/cma6.rds")
cma6_results <- get_cma_results(cma6, march14)
# function output is dataset with summaries for each time interval, regression model, regression summary, and plot for times with NO missings.
cma6_summ_data <- cma6_results$data_summary %>%
  mutate(CMA = 6)
 cma6_plot <- cma6_results$reg_plot + 
 labs(title = "CMA6", x = "Year", y = "Proportion Adherent (80%)") +
  scale_y_continuous(limits = c(0.1,1)) + 
  theme_bw()

cma5 <- readRDS("./data/cma5.rds")
cma5_results <- get_cma_results(cma5, march14)
cma5_summ_data <- cma5_results$data_summary %>%
  mutate(CMA = 5)
cma5_plot <- cma5_results$reg_plot + 
labs(title = "CMA5", x = "Year", y = "Proportion Adherent (80%)") +
  scale_y_continuous(limits = c(0.1,1)) + 
  theme_bw()

cma4 <- readRDS("./data/cma4.rds")
cma4_results <- get_cma_results(cma4, march14)
cma4_summ_data <- cma4_results$data_summary %>%
  mutate(CMA = 4)
cma4_plot <- cma4_results$reg_plot + 
labs(title = "CMA4", x = "Year", y = "Proportion Adherent (80%)") +
  scale_y_continuous(limits = c(0.1,1)) + 
  theme_bw()

#its_plots <- (main_plot / cma6_plot / cma5_plot + cma4_plot)
 #its_plots
cma7_summ_data <- cma7_results$data_summary %>%
  mutate(CMA = 7)

cma_comp <- rbind(cma4_summ_data, cma5_summ_data, cma6_summ_data, cma7_summ_data) %>%
  mutate(CMA = factor(CMA, 
                         levels = c(4, 5, 6, 7),
         label = c("CMA 4", "CMA 5", "CMA 6", "CMA 7")))

cma_scatter <- ggplot(cma_comp, aes(x = window.start, y = prop_adherent, color = CMA)) +
    geom_point() + geom_line(linetype = "dotted") +
  scale_y_continuous(limits = c(0.1,1)) + 
  theme_bw() + scale_color_brewer(palette = "Dark2") + 
  labs(x = "Year", y = "Proportion Adherent (80%)") +
  theme(legend.title=element_blank()) 
cma_scatter
```

While PDC was more appropriate measure compared to MPR, still wanted to compare which CMA to best approximate PDC given package options. Chose CMA7 because it seems to best incorporate observation window (treatment time) in the denominator. This seems confirmed when considering categorization and however measures are calculated. CMA7 has the least proportion missings per time of observation. We can see a decline in the number of missings over the first two years, which is expected as patients with year of diagnosis in 2018 and 2019 should not yet have a PDC in 2018 or 2019.

Additional analyses

```{r}
#| echo: false
#| eval: false
#| label: run-simple-OW
# Running these on 4/4/24 and save as RDS for time
# Interval length & OW only based on dispense end 

# Steps I take to make each OW flexible per person
# 1. getting last dispense date for each ID
ID_dispense <- study_pop %>%
  group_by(StudyID) %>%
    arrange(dispense_dt) %>%
     summarise(first_dispense = min(dispense_dt),
          last_dispense = max(dispense_dt),
          days_supply = last(days_supply)) %>%
    mutate(end_supply = last_dispense + days_supply + days(30)) %>%
  select(StudyID, first_dispense, last_dispense, end_supply)

# 2. add columns back
study_run2 <- study_pop %>%
  inner_join(ID_dispense, by = "StudyID") %>%
  mutate(death_dt = case_when(PATIENTVITAL_STATUS == 0 ~ as.Date(paste(PATIENTDATE_OF_LAST_CONTACT_YYYY, PATIENTDATE_OF_LAST_CONTACT_MM, PATIENTDATE_OF_LAST_CONTACT_DD,
                 sep = "-"), format = "%Y-%m-%d")),
         OW_end = case_when(PATIENTVITAL_STATUS == 0 ~ max(as.Date("2018-01-01",format = "%Y-%m-%d"), death_dt),
                            PATIENTVITAL_STATUS == 1 ~ end_supply),
         OW_duration = as.integer(ceiling((difftime(OW_end, "2018-01-01", units = "weeks"))/4) * 4),
         OW_duration = ifelse(OW_duration < 0, 0, OW_duration))

# 3. loop through
result_list <- c()
for (i in unique(study_run2$StudyID)){
  study_data <- study_run2[study_run2$StudyID == i, ]
  result <- run_CMA(study_data, "CMA7")
  result_list[[i]] <- result
}

cma7_simpleOW <- bind_rows(result_list)
# saveRDS(cma7_simpleOW, "./data/cma7-simpleOW.rds")
```

```{r}
#| echo: false
#| warning: false
#| label: OW-difference
#| fig-cap: "OW without 5-year time window"
cma7_simpleOW <- readRDS("./data/cma7-simpleOW.rds") %>%
  filter(StudyID %in% study_IDs$StudyID)
cma7_OW_results <- get_cma_results(cma7_simpleOW, march14)
# function output is dataset with summaries for each time interval, regression model, regression summary, and plot for times with NO missings.
cma7_OW_plot <- cma7_OW_results$reg_plot 
OW_plot <- cma7_OW_results$reg_plot 
#OW_plot + scale_y_continuous(limits = c(0.1,1))
OW_data <- cma7_OW_results$plot_data
cma7_plot_data <- cma7_results$plot_data

# Now combining to compare between main and shorter OW
OW_data$OW_group <- "Rx end"
cma7_plot_data$OW_group <- "5 year end"
OW_comp <- rbind(OW_data, cma7_plot_data) %>%
  mutate(missing = 8023 - (adherent_count + non_adherent_count),
         OW_group = factor(OW_group, 
                       levels = c('5 year end', 'Rx end'),
         labels = c('Original', 'End of prescription')))

# comparing counts
OW_comp_long <- OW_comp %>%
  pivot_longer(cols = c(adherent_count, non_adherent_count, missing),
               names_to = "fate",
               values_to = "count") %>%
  mutate(Outcome = factor(fate, 
                       levels = c('missing','non_adherent_count', 'adherent_count'),
         labels = c('Missing', 'Not Adherent', 'Adherent'))
         )

ggplot(OW_comp_long) +
    geom_col(aes(x = window.start, y = count, fill = Outcome, color = Outcome)) +
    scale_fill_brewer(palette = "Purples") + 
    scale_color_brewer(palette = "Purples") +
  # TODO edit x axis limits so that year is correct
  #  scale_x_continuous(expand = c(0,0)) + 
  #  scale_y_continuous(expand = c(0,0)) + 
    labs(x = "Year", y = "Count") +
    theme_bw() + facet_grid(~OW_group) +
  theme(legend.title=element_blank())  

# TODO would like to differentiate missings who have not yet started treatment
# may need to get cma aggregates and incorporate that in tallies. 
```

Curious if we were to include those with only one claim: 
```{r}
#| echo: false
#| eval: false
#| label: multiple-claims
#| fig-cap: "claims"

cma7_old <- readRDS("./data/cma7_all.rds")

cma7_old2 <- get_cma_results(cma7_old, march14)

main_plot <- cma7_results$plot_data %>%
  mutate(claims = "2+ claims")
all_claims_plot <- cma7_old2$plot_data %>%
  mutate(claims = "Incl. 1 claim")

claim_compare <- rbind(all_claims_plot, two_claims_plot)

two <- c("black", "#932667")
ggplot(claim_compare, aes(x = window.start, y = prop_adherent, color = claims)) +
    geom_point(size = 2) +
   # geom_vline(xintercept = march14, colour = "#E41A1C") +  
  #scale_color_brewer(palette = "RdGy") +
  scale_color_manual(values = two) +
  scale_y_continuous(limits = c(0.1,1)) + 
  theme_bw() + 
  labs(x = "Year", y = "Proportion Adherent (80%)") +
  theme(legend.title=element_blank())

# If there is no hormone date not avail, perhaps using date of surgery and/or diagnosis to start the 5 year mark. Date of surgery is not avail for everyone.
# Or just exclude those who we do not have hormone start for? OR use date of first fill?
```

```{r}
#| eval: false
#| echo: false
#| label: run-other-intervals

run_CMA_90days <- function(data, CMA_choice){
  # creating these so can change inside of function to be able to loop thru
  FU_start <- as.Date("2017-10-09", format = "%Y-%m-%d") 
  FU_end <- as.Date("2022-01-01", format = "%Y-%m-%d")
  FU_duration <- as.integer(difftime(FU_end, FU_start, units = "weeks"))
  Jan1 <- as.Date("2018-01-01", format = "%Y-%m-%d")
  
  OW_duration <- min(208, first(data$OW_duration), na.rm = TRUE)
  
    cma <- CMA_sliding_window(CMA.to.apply = CMA_choice,
                     data = data,
                     ID.colname = "StudyID",
                     event.date.colname = "dispense_dt",
                     event.duration.colname = "days_supply",
                     event.daily.dose.colname ="daily_dose",
                     medication.class.colname = "AETdrug",
                     carry.only.for.same.medication = FALSE,
                     consider.dosage.change = FALSE,
                     followup.window.start = FU_start,
                     followup.window.duration = FU_duration,
                     followup.window.duration.unit = "weeks",
                     observation.window.start = Jan1,
                     observation.window.duration = OW_duration,
                     observation.window.duration.unit = "weeks",
                     sliding.window.start = 0, 
                     sliding.window.start.unit = "weeks",
                     sliding.window.duration = 4,
                     sliding.window.duration.unit = "weeks",
                     # if window and step are same, there is no overlap
                     sliding.window.step.duration = 4,
                     sliding.window.step.unit = "weeks",
                     date.format = "%Y/%m/%d",
                     parallel.backend = "none",
                     parallel.threads = 1) 
  result <- as_tibble(cma$CMA)
  return(result)
}
# step 2. add columns back
study_run3 <- study_pop %>%
  inner_join(ID_dispense, by = "StudyID") %>%
  mutate(horm_mo = case_when(is.na(CTCRX_SUMM_DATE_HORMONE_MM) ~ 1,
                              TRUE ~ CTCRX_SUMM_DATE_HORMONE_MM),
         horm_dd = case_when(is.na(CTCRX_SUMM_DATE_HORMONE_DD) ~ 1,
                            TRUE ~ CTCRX_SUMM_DATE_HORMONE_DD),
         horm_dt = case_when(!is.na(CTCRX_SUMM_DATE_HORMONE_YYYY) ~ as.Date(
           paste(CTCRX_SUMM_DATE_HORMONE_YYYY,
                 sprintf("%02d", as.numeric(horm_mo)),
                 sprintf("%02d", as.numeric(horm_dd)),
                 sep = "-"), format = "%Y-%m-%d")),
         horm_end_dt = horm_dt + years(5),
         death_dt = case_when(PATIENTVITAL_STATUS == 0 ~ as.Date(paste(PATIENTDATE_OF_LAST_CONTACT_YYYY, PATIENTDATE_OF_LAST_CONTACT_MM, PATIENTDATE_OF_LAST_CONTACT_DD,
                 sep = "-"), format = "%Y-%m-%d")),
         OW_end = case_when(PATIENTVITAL_STATUS == 0 ~ max(as.Date("2018-01-01",format = "%Y-%m-%d"), death_dt),
                            PATIENTVITAL_STATUS == 1 ~ max(end_supply, horm_end_dt, na.rm = TRUE)),
         OW_duration = as.integer(ceiling((difftime(OW_end, "2018-01-01", units = "weeks"))/4) * 4),
         OW_duration = ifelse(OW_duration < 0, 0, OW_duration))

# 3. loop through
result_list <- c()
for (i in unique(study_run3$StudyID)){
  study_data <- study_run3[study_run3$StudyID == i, ]
  result <- run_CMA(study_data, "CMA7")
  result_list[[i]] <- result
}

cma7_90days <- bind_rows(result_list)
```
