---
title: "Adherence Aim"
format: docx
editor: visual
editor_options: 
  chunk_output_type: console
---

## Medication use during a pandemic

Chrystelle Kiang

### Note space

Here goes my attempt at using quarto for some of this dissertation aim text.

```{r}
#| warning: false
#| include: false
library(here) 
library(tidyverse)
library(gtsummary)
library(flextable)
library(ggplot2)

# this is all prescriptions for eligible study population
# these are created in data_creation.R 
pharmstudydata <- readRDS("studypopdata.rds")
#TODO will need to subset further to actual analytic dataset
#TODO drop repeat variables here? 

# subset to drugs of interest
studydata <- pharmstudydata %>%
  filter(drug_group %in% c("AI","TAMOXIFEN"))

```

### Study Population

First, want to get descriptive statistics... Details about study population and prescription information go here.

```{r}
#| echo: false

# one line per patient
studypop <- studydata %>%
  distinct(StudyID, .keep_all = TRUE)

n_studypop <- n_distinct(studydata$StudyID)

table1 <- studypop %>%
  tbl_summary(include = c(sex, race, hispanic, marital_status, CTCAGE_AT_DIAGNOSIS, CTCDATE_OF_DIAGNOSIS_YYYY, laterality, grade, estrogen_receptor, reportingsource))

as_flex_table(table1)
```

### Results

There are `n_studypop` patients who had ER positive tumors, diagnosed stages I to III, and had any prescription starting on January 1, 2018.

#### Descriptive prescription information

```{r}
#| echo: false
#| include: false
# want to look at prescriptions per person
# only keep one drug per patient
drugsbystudyID <- studydata %>%
  group_by(StudyID) %>%
  summarise(unique_drug_type = n_distinct(drug_group))

studyIDs_multiple_drugs <- drugsbystudyID %>%
  filter(unique_drug_type > 1) 

studydata_multiple_drugs <- studydata %>%
  inner_join(studyIDs_multiple_drugs, by = "StudyID")

# TODO save these plots and print later, or remove the include:false
# studydata is subset to AI and TAM only
ggplot(studydata, aes(x = dispense_dt, y = PersonID, group = PersonID, color = drug_name)) +
  geom_point() + geom_line() + theme_classic()

ggplot(studydata, aes(x = dispense_dt, y = PersonID, group = PersonID, color = drug_group)) +
  geom_point() + geom_line() + facet_wrap(vars(CTCRX_SUMM_DT_SURGERY_YYYY)) +
   labs(title = "all patients") + theme_classic()

# subset of patients with BOTH AI and TAM
ggplot(studydata_multiple_drugs, aes(x = dispense_dt, y = PersonID, group = PersonID, color = drug_group)) + geom_point() + geom_line() + 
  labs(title = "patients with both AI and TAM") + theme_classic()

# prescription duration 
ggplot(data = studydata, aes(x = days_supply)) + geom_histogram()

```

-   To do: Plot of prescriptions over time (separate from ITS)
